---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# brio - Basic R Input Output

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/r-lib/brio/branch/master/graph/badge.svg)](https://codecov.io/gh/r-lib/brio?branch=master)
[![R build status](https://github.com/r-lib/brio/workflows/R-CMD-check/badge.svg)](https://github.com/r-lib/brio/actions)
<!-- badges: end -->

Functions to handle basic input output, these functions always
read and write UTF-8 files and provide more explicit control over line endings.

## Reading files

```{r}
library(brio)
write_lines(c("abc", "123"), "my-file")

# Write with windows newlines
write_lines(c("abc", "123"), "my-file-2", eol = "\r\n")

file_line_endings("my-file")

file_line_endings("my-file-2")

read_lines("my-file")

unlink(c("my-file", "my-file-2"))
```

## Benchmarks

Speed is not necessarily a goal of brio, but it does end up being a nice side effect.

```{r}
gen_random <- function(characters, num_lines, min, max) {
  line_lengths <- sample.int(max - min, num_lines, replace = TRUE) + min
  vapply(line_lengths, function(len) paste(sample(characters, len, replace = TRUE), collapse = ""), character(1))
}

set.seed(42)

# generate 1000 random lines between 100-1000 characters long
data <- gen_random(letters, 1000, min = 100, max = 1000)
```

## Writing

Write speeds are basically the same regardless of method, though brio does avoid some extra memory allocations.

```{r}
bench::mark(
  brio::write_lines(data, "benchmark"),
  readr::write_lines(data, "benchmark"),
  base::writeLines(data, "benchmark"),
  check = FALSE
)
```

## Reading

Reading speeds are a decent amount faster with brio, mainly due to larger block sizes and avoidance of extra copies.

```{r}
bench::mark(
  brio::read_lines("benchmark"),
  readr::read_lines("benchmark"),
  base::readLines("benchmark")
)

unlink("benchmark")
```
